Robust Spectroscopic Inference with Imperfect Models

### 2014-08-27 tex/paper.tex
edits on revised Sect 3, 4

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

@@

@@

In this section, we aim to demonstrate the functionality and highlight the flexibility and utility 
of the probabilistic inference methodology described above.  To do that, we illustrate how the 
modeling framework operates for two representative examples.  The first is an elaboration of the 
example shown throughout Section \ref{sec:method}, using a high resolution optical 
($\sim$5000-5400\,\AA) spectrum of the F5 dwarf main sequence star (and transiting exoplanet host) 
WASP-14 \citep{joshi09,torres12}.  This spectrum is interpreted in the context of two different 
synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
$\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
and the most recent incarnation of the {\sc Phoenix} library \citep{husser13}.  The second 
example uses a medium resolution near-infrared ($\sim$2.0-2.4\,$\mu$m) spectrum of the M5 dwarf 
star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
\citep{cushing05,rayner09}.  In this case we also use the {\sc Phoenix} library, since it forms a complete grid for late spectral types.

@@

\citet{torres12} originally acquired these data in an effort to measure the WASP-14 stellar 
properties, and thereby better constrain the parameters of the transiting exoplanet it hosts.  To 
that end, they employed the {\tt MOOG} and {\tt SPC} approaches outlined in Section 
\ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
free parameter or assumed to be fixed (to $\log g = 4.29$), based on an inference of the mean 
stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
evolution models in the H-R diagram \citep{joshi09}.  Only parameter inferences from the latter 
assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
synthesize a suite of individual \ion{Fe}{1} and \ion{Fe}{2} lines in the $\sim$4300--6750\AA\ 
region, whose equivalent widths were compared with the TRES measurements in a simple least-squares 
sense.  This approach suggests that $T_{\rm eff} = 6150\pm75$\,K and $[{\rm Fe/H}] = 
-0.34\pm0.10$\,dex.  The {\tt SPC} analysis relied on a direct comparison between three TRES 
spectral orders covering $\sim$5100-5400\,\AA\ and synthetic spectra from the {\sc CfA/Kurucz} 
library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
i = 3.9\pm0.5$\,km s$^{-1}$.  It is worth noting that these parameter inferences represent an 
average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
the systematic spread between the best-fit parameter values. \comm{This is something we need to 
discuss more with Dave and Willie.}

Given the decision to employ a spectral library to generate models
(Sect.~\ref{subsec:synthetic}), it is most relevant to compare the
\citeauthor{torres12}~{\tt SPC} result with our framework.  Using the same TRES
spectral orders (for only a single WASP-14 observation, as noted above), the
{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log
g$, we infer parameters that are entirely consistent with the {\tt SPC}
approach: $T_{\rm eff} = 6517\pm13$\,K, $[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and
$v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~\ref{fig:Kurucz_posterior}
displays some relevant marginal projections of the posterior-space, and
Table~\ref{table:Kurucz} compiles our estimates of the peaks of the marginal
posteriors and associated confidence intervals for all parameters employed in
the modeling.  Figure~\ref{fig:Kurucz_residuals} demonstrates the overall
quality of this approach in reproducing the WASP-14 spectrum and, most
significantly, the structure of the associated residuals.
Figure~\ref{fig:metacomparison} compares the inferred parameters with the
various determinations made by \citet{torres12}.

\begin{figure}[!htb]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
  \caption{The posterior probability function of the stellar parameters for WASP-14 as explored by the MCMC Gibbs sampler. These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$). Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
}
\label{fig:Kurucz_posterior}
\end{center}
\end{figure}

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_Kurucz_logg.pdf}
\caption{A sample of the WASP-14 data set overlaid with a model spectrum produced by parameters drawn from the posterior probability distribution (fixing $\log_{10} g = 4.29$ dex) and using the {\sc CfA/Kurucz} synthetic library. The residuals are shown zoomed along with the three-sigma contours of random draws from the covariance matrix, after Figure~\ref{fig:matrix}.}
\label{fig:Kurucz_residuals}
\end{center}
\end{figure*}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6517 & 13 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex\tablenotemark{a}} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.62 & 0.02 $\kms$ \\
$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
$\cc{22}{1}$ & -0.014 & 0.002\\
$\cc{22}{2}$ & -0.014 & 0.002\\
$\cc{22}{3}$ & -0.003 & 0.001\\
\\
$b_{22}$ & 1.04 & 0.02 \\
$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
$l_{22}$ & 7.5 & 0.8 $\kms$ \\

\cutinhead{Order 23 nuisance parameters}
$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
$\cc{23}{1}$ & -0.013 & 0.002 \\
$\cc{23}{2}$ & -0.025 & 0.001 \\
$\cc{23}{3}$ &  0.001 &  0.001\\
\\
$b_{23}$ & 1.01 & 0.02 \\
$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
$l_{23}$ & 6.5 & 0.5 $\kms$\\

\cutinhead{Order 24 nuisance parameters}
$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00\tablenotemark{a}} \\
$\cc{24}{1}$ & -0.010 & 0.002 \\
$\cc{24}{2}$ & -0.011 & 0.001\\
$\cc{24}{3}$ &  0.000 & 0.001\\
\\
$b_{24}$ & 1.02 & 0.02 \\
$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
$l_{24}$ & 7.3 & 0.6 $\kms$ \\
\enddata
\tablenotetext{a}{parameter value fixed}
\end{deluxetable}

\begin{figure}[!htb]
\begin{center}
\includegraphics{figs/metacomparison.pdf}
\caption{Comparing all the posteriors.}
\label{fig:metacomparison}
\end{center}
\end{figure}

To understand the apparent differences in the quoted precisions, it is beneficial to review what factors 
into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
range that no local covariance kernels need to be instantiated; in this specific case we use only a 
global covariance kernel.}  To break down their respective contributions, we performed the 
same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
are generated from the model; in this paradigm, the only way to introduce deviation is through 
white noise (see Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is invalid, since just a few simple parameters 
($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
posteriors. For synthetic models which are less accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend to dominate the error budget.

In addition to statistical and systematic error, one might expect that systematic instrumental error might also contribute to the parameter uncertainties. Commonly, one might estimate this error by fitting spectra taken by different instruments or at different times and then averaging the results. Additionally, one may add in an additional uncertainty ``floor'' that maps to the empirical scatter of typical spectroscopic fits. Such procedures generally make the assumption that this additional error contribution is Gaussian and independent in each parameter of interest. Our hierarchical Bayesian model avoids such a heuristic by
parameterizing the quality of fit self-consistently: because the parameter uncertainty is derived through the model itself, it preserves correlations between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$. Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters. However, for well-calibrated data sets from the same instrument any ill-effects are generally small. To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim10$ K and $\lesssim 0.001$ dex in $\Z$. When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.

\comm{Finally, you can transition into the model systematics issue.  Here you want to do the 
modeling for Phoenix in orders 22-24 and demonstrate that you get much different parameters.  You 
will have to show the results, and then can take that opportunity to discuss the local covariance 
kernels.  Just show the results and make these two points, don't belabor the details: I don't 
know what to make of the weird Phoenix parameters.}

\begin{figure}[!htb]
\begin{center}
  \includegraphics[draft, width=0.5\textwidth,height=2in]{figs/PHOENIX_triangle.pdf}
\caption{ {\sc Phoenix} WASP-14 marginal posteriors.}
\label{fig:PHOENIX_posterior}
\end{center}
\end{figure}

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_PHOENIX_logg.pdf}
\caption{ {\sc Phoenix} WASP-14 model and residuals.}
\label{fig:residuals_PHOENIX}
\end{center}
\end{figure*}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6516.8 & 12.6 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.62 & 0.02 $\kms$ \\
$\log_{10} \Omega$ & -12.72 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\cc{22}{0}$ & & \\
$\cc{22}{1}$ & & \\
$\cc{22}{2}$ & & \\
$\cc{22}{3}$ & & \\
$b_{22}$ & & \\
$a_{{\rm g}, 22}$ & & \\
$l_{22}$ & & \\
\cutinhead{Order 23 nuisance parameters}
$\cc{23}{0}$ & & \\
$\cc{23}{1}$ & & \\
$\cc{23}{2}$ & & \\
$\cc{23}{3}$ & & \\
$b_{23}$ & & \\
$a_{{\rm g}, 23}$ & & \\
$l_{23}$ & & \\
\cutinhead{Order 24 nuisance parameters}
$\cc{24}{0}$ & & \\
$\cc{24}{1}$ & & \\
$\cc{24}{2}$ & & \\
$\cc{24}{3}$ & & \\
$b_{24}$ & & \\
$a_{{\rm g}, 24}$ & & \\
$l_{24}$ & & \\
\enddata
\end{deluxetable}

\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
orders with Phoenix.}

While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range with no specific tuning for solar type stars, there are likely to be a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8). Some routines choose to ``sigma-clip'' the outlier lines and procede with fitting, however with such an approach it is difficult to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}. For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?

This behaviour motivates the global and local kernels in a more sophisticated likelihood function, which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.  The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum. Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the outlier lines.

Using the {\sc Phoenix} models on WASP-14, we infer substantially different posteriors than with using the {\sc CfA/Kurucz} library. Figure~\ref{fig:PHOENIX_posterior} displays some of the relevant marginal posteriors, and Table~\ref{table:PHOENIX} compiles the estimates of the peaks of the posterior distribution and associated confidence intervals. Figure~\ref{fig:residuals_PHOENIX} demonstrates the quality of this approach and the structure of the associated residuals. Compared to the Kurucz residuals in Figure~\ref{fig:residuals_Kurucz}, there are two significant differences. First, the amplitude of the global covariance structure is larger. The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance. Second, local covariance kernels track outlier spectral lines. These kernels can map out the regions of the spectrum which are outliers. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner. The locations and amplitudes of these kernels output from the code would be useful information to consider when refining the models of stellar sythenis.

Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveals that there is a systematic offset depending on which spectral library is used as a backend. Depending on the choice of spectral library, there may be systematic differences of 200 K or more in $T_{\rm eff}$ and 0.2 dex or more in $\Z$.

@@

A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
(IRTF).  SPEX is a cross-dispersed echelle that covers the red-optical to thermal-infrared spectrum 
(0.7--5.5\,$\mu$m) in two spectrograph settings.  These data were obtained as part of the IRTF 
spectral standard library project \citep{cushing05,rayner09}, and were processed through the 
well-vetted {\tt Spextool} pipeline \citep{cushing04,vacca03} to deliver a fully-calibrated 
spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.

>>>>

%\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\title{Robust Spectroscopic Inference with Imperfect Models}
\author{Ian Czekala, Sean M.~Andrews, Guillermo Torres, \& David W.~Latham}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

@@

\begin{figure}[!htb]
\begin{center}
  \includegraphics{figs/badlines0.pdf}
  \includegraphics{figs/badlines1.pdf}
  \includegraphics{figs/badlines2.pdf}
  \includegraphics{figs/badlines3.pdf}
  \caption{A collection of spectral lines which have imperfect model fits.
    ( {\it first panel}) The majority of spectral lines ($\gtrsim 60$\%) will
    have minor differences in strength between data and model spectrum, which
    produce low-amplitude correlations in the residuals on the length scale of
    the width of a typical spectral line.  ( {\it second panel}) Sometimes
    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
    patch of negative residuals.  ( {\it third panel}) Sometimes ($\lesssim
    5$\% of all lines), an extraneous line in the model leaves a large, highly
    correlated patch of positive residuals. ({\it fourth panel}) If the line
    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
    there will be many correlated residuals of moderate amplitude. For any specific line, there might
    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
    will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure}

@@

In this section, we aim to demonstrate the functionality and highlight the flexibility and utility 
of the probabilistic inference methodology described above.  To do that, we illustrate how the 
modeling framework operates for two representative examples.  The first is an elaboration of the 
example shown throughout Section \ref{sec:method}, using a high resolution optical 
($\sim$5000-5400\,\AA) spectrum of the F5 dwarf main sequence star (and transiting exoplanet host) 
WASP-14 \citep{joshi09,torres12}.  This spectrum is interpreted in the context of two different 
synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
$\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
and the most recent incarnation of the {\sc Phoenix} library \citep{husser13}.  The second 
example uses a medium resolution near-infrared ($\sim$2.0-2.4\,$\mu$m) spectrum of the M5 dwarf 
star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
\citep{cushing05,rayner09}.  In this case we use the {\sc Phoenix} library to generate models, 
since it is well-sampled at cooler temperatures.

@@

\citet{torres12} originally acquired these data in an effort to measure the WASP-14 stellar 
properties, and thereby better constrain the parameters of the transiting exoplanet it hosts.  To 
that end, they employed the {\tt MOOG} and {\tt SPC} approaches outlined in Section 
\ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
free parameter or assumed to be fixed (to $\log g = 4.29$) based on an inference of the mean 
stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
evolution models in the H-R diagram \citep{joshi09}.  Only parameter inferences from the latter 
assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
synthesize a suite of individual \ion{Fe}{1} and \ion{Fe}{2} lines in the $\sim$4300--6750\AA\ 
region, whose equivalent widths were compared with the TRES measurements in a simple least-squares 
sense.  This approach suggests that $T_{\rm eff} = 6150\pm75$\,K and $[{\rm Fe/H}] = 
-0.34\pm0.10$\,dex.  The {\tt SPC} analysis relied on a direct comparison between three TRES 
spectral orders covering $\sim$5100-5400\,\AA\ and synthetic spectra from the {\sc CfA/Kurucz} 
library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
i = 3.9\pm0.5$\,km s$^{-1}$.  It is worth noting that these parameter inferences represent an 
average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
the systematic spread between the best-fit parameter values. \comm{This is something we need to 
discuss more with Dave and Willie.}

Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
a direct comparison of the \citeauthor{torres12}~{\tt SPC} result with our framework is relevant.  
Using the same TRES spectral orders (for a single WASP-14 observation, as noted above), the {\sc 
CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer parameters 
that are entirely consistent with the {\tt SPC} approach; see Table~\ref{table:Kurucz}.  
Figure~\ref{fig:Kurucz_residuals} demonstrates the quality of our approach in reproducing the 
WASP-14 spectrum and, especially significant, the structure of the residuals.  
Figure~\ref{fig:Kurucz_posterior} shows some marginal projections of the posterior-space, 
highlighting key parameter degeneracies.  
%Figure~\ref{fig:metacomparison} compares the inferred 
%parameters with the various determinations made by \citet{torres12}.

\begin{figure*}[!t]
\includegraphics{figs/residuals_Kurucz_logg.pdf}
\figcaption{({\it top}) A representative segment of the TRES spectrum of WASP-14 ({\it blue}), 
overlaid with a {\sc CfA/Kurucz} model ({\it red}) generated by drawing parameters from the 
posterior distribution (for a fixed $\log g = 4.29$).  ({\it bottom}) The residual spectrum along 
with contours representing the distributions of a large number of random draws from the covariance 
matrix (the shading is representative of the 1, 2, and 3\,$\sigma$ spreads of that distribution of 
draws), as in Fig.~\ref{fig:matrix}.  With such a finely tuned model library, local covariance 
kernels are not required.  \label{fig:Kurucz_residuals}}
\end{figure*}

\begin{deluxetable}{lr@{ $\pm$ }l|r@{ $\pm$ }l}[!b]
\tablecaption{\label{table:Kurucz}Inferred Parameters for WASP-14}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc CfA/Kurucz}} & \multicolumn{2}{c}{{\sc Phoenix}}}
\startdata
$T_{\rm eff}$ (K) & 6517     & 13                      &  6273   & 15        \\
$\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
$\Z$          & -0.27    & 0.01                    & -0.498  & 0.006     \\
$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.29     & 0.05                    &  4.89   & 0.05      \\
$v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
$\log \Omega$\tablenotemark{b} & -12.7181 & 0.0004 & -19.678 & 0.001
\enddata
\tablenotetext{a}{Small differences in the rotational broadening between the two libraries are
expected, since they assume different micro- and macro-turbulence behavior in the stellar
atmospheres.}
\tablenotetext{b}{The {\sc CfA/Kurucz} spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.  The inference on $\Omega$ from the {\sc Phoenix} library is in steradian units.}
\tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
distributions for each parameter.  The quoted uncertainties correspond to the 68.3\%\ 
($\sim$1\,$\sigma$) confidence intervals.  Expanded lists of the inferred values and uncertainties 
for all nuisance parameters are available electronically.}
\end{deluxetable}

\begin{figure}[!b]
  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
  \figcaption{The posterior probability distributions for the interesting stellar parameters 
(marginalized over all other model parameters) of WASP-14 based on the {\sc CfA/Kurucz} model 
library, as explored by the MCMC Gibbs sampler.  For the two-dimensional posterior projections, 
contours are drawn at the standard 1, 2, and 3\,$\sigma$ levels (of a normal distribution) for 
reference.  \label{fig:Kurucz_posterior} }
\end{figure}

%\begin{figure}[!htb]
%\begin{center}
%\includegraphics{figs/metacomparison.pdf}
%\caption{The posterior peaks and widths for various determination of the stellar parameters of WASP-14. The determinations in \citet{torres12} are shown along with the estimates using the {\sc CfA/Kurucz} synthetic library and the {\sc PHOENIX} synthetic library. \todo{What else do we want to show on here?}}
%\label{fig:metacomparison}
%\end{center}
%\end{figure}

It is beneficial to review what factors into the parameter uncertainties.  In our approach, we 
consider contributions from both trivial (Poisson) noise and a non-trivial Gaussian process 
covariance matrix.\footnote{Note that the {\sc CfA/Kurucz} library is so well-tuned for stars like 
WASP-14 over this limited spectral range that no local covariance kernels need to be instantiated; 
in this specific case, we employ only global covariance kernels to reproduce the residual 
structure.} To decompose their respective contributions, we performed the same analysis with only a 
trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the likelihood function to the 
familiar $\chi^2$ form.  Doing this, we identify little bias to the best-fit parameters, but find 
unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, $[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v 
\sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small uncertainties are to be expected, since the 
prescribed likelihood function presumes that the data are generated from the model; in this 
paradigm, the only way to introduce deviation is through white noise (see Fig.~\ref{fig:matrix}).  
Of course, that implicit assumption is invalid, since just a few simple parameters ($\vT$) are 
generally insufficient to account for the thousands of pixels and hundreds of spectral features in 
a typical dataset.  In essence, the data and an imperfect model spectrum also have {\it systematic} 
differences that also contribute to the uncertainties, which we forward-model using a non-trivial 
covariance matrix.  In this specific case, these systematic and the Poisson uncertainties 
contribute roughly equally to the $\vT$ posteriors.  For synthetic models that are less finely 
tuned than the {\sc CfA/Kurucz} grid, the systematic uncertainties will dominate (see below).

In addition to noise in the data (statistical uncertainties) and systematic uncertainties in the 
models, we might also expect that systematics associated with the instrument and/or calibration 
contribute to the parameter uncertainties.  Common practice is to estimate these latter systematics 
by comparing parameter inferences made using different spectra (e.g., from different instruments or 
observing nights): the results are usually presented as an average, with uncertainties inflated by 
adding in a ``floor" term that accounts for the dispersion between the inferences from individual 
spectra.  That procedure implicitly assumes that these systematics are Gaussian (add in quadrature) 
and independent in each parameter.  But the approach we are advocating here avoids such a heuristic 
by self-consistently parameterizing the fit quality.  Because the parameter uncertainties are 
derived through a forward model, this method (crucially) preserves the morphologies of intrinsic 
parameter degeneracies in a way that cannot be reproduced by artificially (i.e., manually) 
inflating the posteriors.

To quantify these instrumental systematics for WASP-14, we modeled the three other TRES spectra 
obtained by \citet{torres12} independently (under the same assumptions).  We found a minimal 
scatter of $\sim$20\,K in $T_{\rm eff}$ and $\sim$0.02\,dex in $\Z$ between the individual 
inferences, a testament to quality calibration and an exceptionally stable spectrograph.  The more 
appropriate way of combining the inferences is to model these four spectra simultaneously in our 
hierarchical framework: in that case, we found that the parameter inferences become slightly more 
precise (as one would expect for a weighted average and well-tuned models), while preserving the 
intrinsic correlations between parameters (something expressly destroyed in a weighted average).

%\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
%orders with Phoenix.}

While comparisons with the {\sc CfA/Kurucz} models confirm that our methodology can faithfully 
reproduce the inferences made with standard spectral fitting routines, they do little to illustrate 
a key feature -- the treatment of local patches of large covariance (see 
Sect.~\ref{subsec:local_covariance}) in a more sophisticated likelihood function.  For that we turn 
to the {\sc Phoenix} library, comprised of model spectra with a much broader spectral range that 
have not been fine-tuned for solar type stars.  As a result, users should expect that there will 
likely be a population of ``outlier" residual lines present when comparing any observed spectrum 
with the {\sc Phoenix} models \citep[e.g.,][their Fig.~8]{husser13}.  The standard approach is to 
outright ignore, or perhaps ``$\sigma$-clip", the outlier lines and proceed with fitting the 
remainder of the spectrum.  However, it is difficult to quantify how such decisions ultimately 
affect the parameter inferences \citep{hogg10}.  By parameterizing these residual features with 
local covariance kernels, we are self-consistently preserving any discriminatory power against the 
stellar parameters that might still be contained in their imperfect representations.

\begin{figure*}[!htb]
\includegraphics{figs/residuals_PHOENIX_logg.pdf}
\figcaption{Same as Fig.~\ref{fig:Kurucz_residuals} for the {\sc Phoenix} models.  Note the 
increased vertical scale and local covariance structure for the residuals.
 \label{fig:PHOENIX_residuals}}
\end{figure*}

\begin{figure}[!b]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/PHOENIX_triangle.pdf}
\figcaption{Same as Fig.~\ref{fig:Kurucz_posterior}, but for the {\sc Phoenix} models.  
\label{fig:PHOENIX_posterior}}
\end{center}
\end{figure}

%This behavior of the {\sc PHOENIX} library motivates the use of global and local covariance kernels in a sophisticated likelihood function. The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum, and acts as a mechanism to account for the systematic mismatch of data spectrum and model spectrum. Including this covariance broadens the posteriors on the stellar parameters, while the local covariance kernels provide cushion against outlier lines.

Figure~\ref{fig:PHOENIX_residuals} illustrates the approach in practice through a comparison of the 
WASP-14 data and the {\sc Phoenix} models.  There are two notable differences in the residual 
spectra presented there and in the corresponding Figure~\ref{fig:Kurucz_residuals} for the {\sc 
CfA/Kurucz} models.  First, the amplitude of the global covariance structure is $\sim$2$\times$ 
larger for the {\sc Phoenix} models: since that amplitude is an indirect parameterization of the 
fit quality, this indicates that the {\sc Phoenix} models present an overall worse match to the 
data.  Second, there are several local regions of strong residuals (outlier lines), which are 
enveloped in a parametric treatment of the local covariance structure.  In effect, we have modeled 
the large residuals with a suite of individual covariance kernels, with parameters that identify 
and appropriately down-weight imperfect model spectral lines in a self-consistent manner.

Aside from the structure of the covariance matrix, there is also a striking difference between the 
posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
CfA/Kurucz} model libraries: see Table~\ref{table:Kurucz} and compare 
Figures~\ref{fig:Kurucz_posterior} and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find 
a systematically cooler temperature (by $\sim$240\,K) and lower metallicity (by a factor of 
$\sim$2) than inferred from the {\sc CfA/Kurucz} models.  Because our framework explicitly 
mitigates the potential for posterior bias from imperfect matches between data and models, we can 
safely conclude that these apparent discrepancies are intrinsic to the different input physics of 
the model libraries.\footnote{It is worth noting that these differences are not a manifestation of 
the stringent prior on surface gravity.  When the spectra are modeled with $\log g$ unconstrained, 
the posteriors for both libraries shift, but their respective discrepancies actually increase 
slightly.}  Such a comparison serves as a stark reminder that incomplete and/or over-simplified 
physical prescriptions remain the most significant source of systematic uncertainty in estimates 
of fundamental stellar parameters.

@@

A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
(IRTF).  SPEX is a cross-dispersed echelle spectrograph that covers the red-optical to 
thermal-infrared spectrum (0.7--5.5\,$\mu$m) in two settings.  These data were obtained as part of 
the IRTF spectral standard library project \citep{cushing05,rayner09}, and were processed through 
the well-vetted {\tt Spextool} reduction pipeline \citep{cushing04,vacca03} to deliver a fully 
calibrated spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.

Modeling late-type stellar atmosphere structures and their spectra is considerably more complex 
than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
opacities.  Especially confounding is the presence of complex condensates (clouds) at the coolest 
temperatures \citep{allard13}, making it considerably more challenging to determine (sub-)stellar 
properties \citep{rajpurohit14}.  Various approaches have been taken to infer stellar parameters in 
the face of these difficulties, including iteratively masking regions with poor spectral agreement 
\citep[e.g.,][]{mann13}.  Astutely, \citeauthor{mann13}~note that such a scheme may exclude 
regions of the spectrum that contain intrinsically useful information for discriminating between 
physical properties, and that a more sophisticated approach would weight each spectral region based 
on its consistency with the data.  The modeling framework that we have constructed here does 
exactly that.

\begin{figure*}[!htb]
  \includegraphics{figs/residuals_Gl51_logg.pdf}
  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a {\sc Phoenix} model 
({\it red}) generated by drawing parameters from the posterior distribution (for a fixed $\log g = 
5.0$).  ({\it bottom}) The residual spectrum along with contours representing the distributions of 
a large number of random draws from the covariance matrix (the shading is representative of the 1, 
2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:}.  Note how the 
``outlier" features are clearly identified and treated by the local covariance kernels.
\label{fig:Gl51_residuals}}
\end{figure*}
